---
title: "PS2 Weiner"
author: "Anna Weiner"
date: "2024-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align='center', message=FALSE, warning=FALSE)
library(tidyverse)
library(knitr)
```

Problem 1: https://github.com/annaweiner2838/PSet2.git

Problem 2: 
```{r}
  library(tidyverse)

# open my data
  gspace = read_csv('greenspace_data_share.csv') #include quotations

# summarize average urban greenspace by region
  table <- gspace |> #made table an object with <- and put on same line
    group_by(Major_Geo_Region) |> #add piping in
    summarise(
      obs = n(), #added comma
      avg = mean(annual_avg_2020), #added _
      weighted_avg = mean(annual_weight_avg_2020) #added _
    ) 

# output as table
  kable(table) #changed to name of object
```

Problem 3: The green space data covers `r length(gspace$City)` urban areas.

Problem 4: In a couple of sentences and with reference to a well-formatted tabulation, describe the green space classification scores for urban areas in 2021.

```{r}
gsscore <- gspace |>
  count(
      Exceptionally_High = peak_NDVI_2021 >= 0.7,
      Very_High = 0.6 <= peak_NDVI_2021 & peak_NDVI_2021 < 0.7,
      High = 0.5 <= peak_NDVI_2021 & peak_NDVI_2021,
      Moderate = 0.4 <= peak_NDVI_2021 & peak_NDVI_2021 < 0.5,
      Low = 0.3 <= peak_NDVI_2021 & peak_NDVI_2021 < 0.4,
      Very_Low = 0.2 <= peak_NDVI_2021 & peak_NDVI_2021 < 0.3,
      Exceptionally_Low = peak_NDVI_2021 < 0.2
  )
col.names = c('Exceptionally_High', 'Very_High', 'High', 'Moderate', 'Low', 'Very_Low', 'Exceptionally_Low')
table(col.names, gsscore$n) ##outputs weird but its CLOSE
```

Problem 5a: a. In 2015 `r sum(gspace$peak_NDVI_2015 >= 0.5)` cities scored `High` or above.  
  
```{r}
print(sum(gspace$peak_NDVI_2015 >= 0.5))
```

Problem 5b: Scored `Exceptionally Low` at any point in the years covered.  
```{r}

```

Problem 5c: Urban areas in arid climate that became greener (as measured by annual weighted average) from 2010 to 2020.  
```{r}

```

Problem 6: How many urban areas became less green (measured by annual average) from 2010 to 2021? Were these changes concentrated in a particular geographic or climate region? Explain (with evidence, of course).

Problem 7: Present a histogram showing the change in greenspace (annual average) from 2010 to 2021. Note that you will need to create a new variable equal to this difference.  

Problem 8: Present a scatter plot of population weighted greenspace in 2021 over the greenspace in 2010. 

**BONUS OPPORTUNITY**: Use color-coding to differentiate urban areas that added versus lost greenspace in that time. Then include a 45-degree line to further highlight the change.